---
title: "Afrigen Kenya Scholars Analysis Notebook"
author: "Bonface Onyango"
date: "`r Sys.Date()`"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Setup and Libraries

# Activate r_env

-   Libraries installed in virtual r_env

<!-- -->

-   conda activate r_env

```{r}
# Core libraries 
library(dplyr) 
library(ggplot2) 
library(igraph) 
library(ggraph) 
library(readr)
library(devtools)
```

## Intall lisearch

```{r}
#uncomment or install in r env
#library(devtools)      # now should work
#install_github("elizagrames/litsearchr", ref="main")
```

```{r}
library(litsearchr)
```

```{r}
packageVersion("litsearchr")
```

# Pubmed search page

-   <https://pubmed.ncbi.nlm.nih.gov/advanced/>

-   search terms (medication) AND (CBT) AND (phobia)

-   Query: ((medication) AND (CBT)) AND (phobia)

```{r}
df_pubmed<-import_results(file="../pubmed-breastcanc-set.nbib")
df_pubmed <- as.data.frame(df_pubmed)
head(df_pubmed)
```

```{r}
nrow(df_pubmed)
```

```{r}
#We can take a look at the first few.
df_pubmed[1:3,]
```

```{r}
# column names
colnames(df_pubmed)
```

```{r}
class(df_pubmed[1, "title"])
```

```{r}
df_pubmed[1, "title"]
```

```{r}
# KEYWORDS
# SOMme arrticles do not contain keywords like article 1
```

```{r}
df_pubmed[1, "keywords"]
```

```{r}
df_pubmed[2, "keywords"]
```

```{r}
# how many are missing keywords
# NULL → absence of a value or empty object → is.na() does not detect it.
is.null(df_pubmed[["keywords"]])     # checks if entire column is NULL
```

# Check null values -\>193

```{r}
sum(sapply(df_pubmed$keywords, is.null))

```

# Extract keywords-\>tagged

```{r}
extract_terms(keywords=df_pubmed[, "keywords"], method="tagged")
```

# Extract keywords

```{r}
library(dplyr)
library(purrr)
library(dplyr)
library(purrr)
library(stringr)

clean_keywords <- function(x) {
  # Handle NULL or empty
  if (is.null(x) || length(x) == 0) return(character(0))
  
  # Remove NA
  x <- x[!is.na(x)]
  if (length(x) == 0) return(character(0))
  
  # Clean and split
  x %>%
    str_replace_all("\\n", " ") %>%     # remove newlines
    str_replace_all("\"", "") %>%       # remove quotes
    str_trim() %>%                      # trim whitespace
    str_split(",") %>%                  # split by comma
    unlist() %>%                        # flatten
    str_trim() %>%                      # trim again
    discard(~ .x %in% c("", "NULL", "character0"))
}

# Apply to df_pubmed$keywords and flatten
df_pubmed <- df_pubmed %>%
  mutate(keywords_clean = map(keywords, clean_keywords))

# Flatten all keywords
keywords_flat <- df_pubmed$keywords_clean %>%
  flatten_chr()

# Remove empty
keywords_flat <- keywords_flat[keywords_flat != ""]
keywords_flat[1:5]
# Preview
```

## Tagged keywords

```{r}
keywords_extracted <- extract_terms(
  keywords = keywords_flat,
  method = "tagged"
)

# --- Step 4: Preview ---
keywords_extracted[1:20]
```

## Tagged keywords -\> min1

```{r}
keywords_extracted_min_1 <- extract_terms(
  keywords = keywords_flat,
  method = "tagged",
  min_n = 1
)

# --- Step 4: Preview ---
keywords_extracted_min_1[1:20]
```

```{r}
keywords_extracted <- extract_terms(
  keywords = keywords_flat,
  method = "tagged",
  min_n = 2
)

print(length(keywords_extracted))
keywords_extracted[1:20]
```

# Tittle search -\>105

```{r}
extract_terms(text=df_pubmed[, "title"], method="fakerake", min_freq=3,
min_n=2)
```

# Stop words

```{r}
clinpsy_stopwords <- read_lines("../clin_psy_stopwords.txt")
clinpsy_stopwords
```

```{r}
all_stopwords <- c(get_stopwords("English"), clinpsy_stopwords)
all_stopwords
```

# Keywords from tittle after removing stop words -\>78

```{r}
title_terms <- extract_terms(
text=df_pubmed[, "title"],
method="fakerake",
min_freq=3, min_n=2,
stopwords=all_stopwords
)
title_terms
```

## Keywords + title terms

```{r}
terms <- unique(c(keywords_extracted_min_1, title_terms))
terms
```

```{r}
docs <- paste(df_pubmed[, "title"], df_pubmed[, "abstract"])
docs[1]
```

```{r}
dfm <- create_dfm(elements=docs, features=terms)
dfm[1:5, 1:7]
```

```{r}
g <- create_network(dfm, min_studies=3)
```

```{r}
ggraph(g, layout="stress") +
coord_fixed() +
expand_limits(x=c(-3, 3)) +
geom_edge_link(aes(alpha=weight)) +
geom_node_point(shape="circle filled", fill="white") +
geom_node_text(aes(label=name), hjust="outward", check_overlap=TRUE) +
guides(edge_alpha=FALSE)
```

```{r}
strengths <- strength(g)
data.frame(term=names(strengths), strength=strengths, row.names=NULL) %>%
mutate(rank=rank(strength, ties.method="min")) %>%
arrange(strength) ->
term_strengths
term_strengths
```

```{r}
cutoff_fig <- ggplot(term_strengths, aes(x=rank, y=strength, label=term)) +
geom_line() +
geom_point() +
geom_text(data=filter(term_strengths, rank>5), hjust="right", nudge_y=20,
check_overlap=TRUE)
cutoff_fig
```

```{r}
cutoff_cum <- find_cutoff(g, method="cumulative", percent=0.8)
cutoff_cum
```

```{r}
cutoff_fig +
geom_hline(yintercept=cutoff_cum, linetype="dashed")
```

```{r}
get_keywords(reduce_graph(g, cutoff_cum))
```

```{r}
cutoff_change <- find_cutoff(g, method="changepoint", knot_num=3)
cutoff_change

```

```{r}
cutoff_fig +
geom_hline(yintercept=cutoff_change, linetype="dashed")
```

```{r}
g_redux <- reduce_graph(g, cutoff_change[1])
selected_terms <- get_keywords(g_redux)
selected_terms
```

```{r}
extra_terms <- c(
"medication",
"cognitive-behavioural therapy",
"cognitive behavioural therapy"
)
selected_terms <- c(selected_terms, extra_terms)
selected_terms
```

# 
